---
title: "mfa_analysis"
author: "Suyuan Liu, Márton Sóskuthy"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(reshape2)
library(lmerTest)

mfa <- read_csv("mandarin_mfa_output.csv") %>%
  arrange(speaker, file, phone_start)
manual <- read_csv("mandarin_manual_suyuan_output.csv") %>%
  arrange(speaker, file, phone_start)
manual_sijia <- read_csv("mandarin_manual_sijia_output.csv") %>%
  arrange(speaker, file, phone_start)

range01 <- function(x){(x-min(x))/(max(x)-min(x))}
source("levenshtein.r")
```

## Data cleaning

```{r data cleaning}
## combine two datasets
## we'll use alignment based on edit distance - so none of the indexing
## magic is necessary!

dfs_mfa <- list()
dfs_sijia <- list()

for (spk in unique(mfa$speaker)) {
  mfa_chunk <- filter(mfa, speaker==spk)
  suyuan_chunk <- filter(manual, speaker==spk)
  sijia_chunk <- filter(manual_sijia, speaker==spk)
  alignment_mfa <- levenshtein_align(suyuan_chunk$phone, mfa_chunk$phone)
  suyuan_chunk <- suyuan_chunk %>%
    mutate(
      phone_start_suyuan=phone_start,
      phone_start_other=mfa_chunk$phone_start[alignment_mfa],
      other_type="mfa",
      diff=abs(phone_start_other - phone_start_suyuan),
    )
  if (nrow(sijia_chunk > 0)) {
    alignment_sijia <- levenshtein_align(suyuan_chunk$phone, sijia_chunk$phone)
    suyuan_chunk2 <- suyuan_chunk %>%
    mutate(
      phone_start_other=sijia_chunk$phone_start[alignment_sijia],
      other_type="sijia",
      diff=abs(phone_start_other - phone_start_suyuan)
    )
    dfs_sijia[[length(dfs_sijia) + 1]] <- suyuan_chunk2
  }
  dfs_mfa[[length(dfs_mfa) + 1]] <- suyuan_chunk
}

df_1 <- bind_rows(dfs_mfa)
df_2 <- bind_rows(dfs_sijia)
df <- bind_rows(df_1, df_2) %>%
  filter(!is.na(diff))

df <- df[!grepl("[a-zA-Z]", df$word),]

# add in speaker information
df <- df %>%
    mutate(sex = case_when(speaker %in% c("ma_4336","ma_5530","ma_5542","ma_5716","ma_5779","ma_5800", "ma_5906","ma_5975") ~ "f",
                           speaker %in% c("maa_4426","ma_5260","ma_5560","ma_5824")  ~ "m"),
           res = case_when(speaker %in% c("ma_4336", "ma_5800") ~ "Canto",
                           speaker %in% c("ma_5542", "ma_5824") ~ "Shanghai",
                           speaker %in% c("ma_5560", "ma_5716") ~ "Beijing",
                           speaker %in% c("ma_5975", "ma_5779") ~ "Tianjin",
                           speaker %in% c("ma_4426", "ma_5530") ~ "Nanjing",
                           speaker %in% c("ma_5260", "ma_5906") ~ "Xinan"),
           simple_phone = gsub("[0-9]", "", df$phone),
           phone_category = case_when(simple_phone %in% c("a", "ai", "ao", 
                                                   "e", "ei", "i", 
                                                   "ia", "iao", "ie", "ii", "io", "iou", "iu", 
                                                   "o", "ou", 
                                                   "u", "ua", "uai", "ue", "uei", "uo", 
                                                   "v", "va", "ve") ~ "vowel", #25
                                      phone %in% c("b", "d", "g", "k", "p", "t") ~ "stop", #6
                                      phone %in% c("c", "ch", "f", "h", "s", "sh", "x", "z", "zh", "q", "j") ~ "fricative", #9
                                      phone %in% c("l", "r") ~ "approximant", #2
                                      phone %in% c("m", "n", "ng") ~ "nasal", #3
                                      )) 

df$res <- factor(df$res,
                 levels=c(
                   "Beijing",
                   "Canto",
                   "Shanghai",
                   "Nanjing",
                   "Tianjin",
                   "Xinan"
                 ))
df$comparison_type <- ifelse(
  df$other_type=="mfa",
  "mfa vs. manual 1",
  "manual 1 vs. manual 2"
) %>%
  factor(levels=c("manual 1 vs. manual 2",
                  "mfa vs. manual 1"))
```

## Creating combined plots

By-speaker.

```{r plot violin plot by speaker}



df_means <- df %>%
  group_by(comparison_type, speaker, res) %>%
  summarise(diff_mean=mean(diff),
            diff_median=median(diff),
            mean_symbol="μ",
            diff_mean_label=paste0(round(diff_mean*1000), " ms"),
            diff_median_label=paste0(round(diff_median*1000), " ms")) %>%
  ungroup()

mean_fun <- function(x){
  print(x)
  return(data.frame(y = x, label = paste0("  ", x, " ms")))
}

ggplot(df, aes(y=diff*1000, x=res, color=res, shade=speaker)) +
  facet_grid(.~comparison_type, drop=T, scales="free_x", space="free_x") +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=res), col=NA, alpha=0.15) +
  geom_text(data=df_means, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9)) +
  geom_text(data=df_means, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9)) +
  geom_point(data=df_means, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9),
             shape=21, fill=NA) +
  geom_point(data=df_means, aes(y=diff_median*1000), size=1, position=position_dodge(width=0.9),
             shape=16, fill=NA) +
  geom_text(data=df_means, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9)) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000, 3000),
                limits=c(0.5, 3500)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank(),
        strip.text=element_text(size=14, colour="black"),
        strip.background=element_blank())

ggsave(filename = "asa_graphs/Mandarin_by_speaker.png",  width = 13, height = 5, bg = "transparent")

```

By-variety.

```{r plot violin plot by variety}
df_means <- df %>%
  filter(other_type=="mfa") %>%
  group_by(res) %>%
  summarise(diff_mean=mean(diff),
            diff_median=median(diff),
            mean_symbol="μ",
            diff_mean_label=paste0(round(diff_mean*1000), " ms"),
            diff_median_label=paste0(round(diff_median*1000), " ms")) %>%
  ungroup()
  

mean_fun <- function(x){
  print(x)
  return(data.frame(y = x, label = paste0("  ", x, " ms")))
}

ggplot(filter(df, other_type=="mfa"), aes(y=diff*1000, x=res, color=res)) +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=res), col=NA, alpha=0.15) +
  geom_text(data=df_means, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9)) +
  geom_text(data=df_means, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9)) +
  geom_point(data=df_means, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9),
             shape=21, fill=NA) +
  geom_point(data=df_means, aes(y=diff_median*1000), size=1, position=position_dodge(width=0.9),
             shape=16, fill=NA) +
  geom_text(data=df_means, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9)) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000, 3000),
                limits=c(0.5, 3500)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank())

ggsave(filename = "asa_graphs/Mandarin_by_variety.png",  width = 8.933, height = 5, bg = "transparent")

```

## For mfa vs. manual only

```{r consonant vs. vowel}
df_mfa <- filter(df, other_type=="mfa")

p_c_v <- ggplot(df_mfa, aes(y=diff*1000, x=phone_category, color=phone_category)) +
  #scale_color_manual(values = c("dodgerblue2", "darkgoldenrod2", 
  #                              "coral2", "coral2",
  #                              "dodgerblue2","darkgoldenrod2",
  #                              "seagreen", "seashell",
  #                              "firebrick1", "firebrick4",
  #                              "mediumorchid3", "mediumorchid4")) +
  #facet_wrap(~phone_category) +
  coord_trans(y = 'log') +
  scale_y_continuous(breaks=c(10,50, 100, 500, 1000, 2000, 3000)) +
  # using violins to show distributions per vowel
  geom_boxplot() +
  # adding the median separately as a point
  theme_bw() + 
  labs(x = "Variation of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none") +
  theme(text=element_text(size=15))

ggsave(p_c_v, filename = "asa_graphs/Mandarin_consonants_vowel.png",  width = 6, height = 5, bg = "transparent")

```

```{r plot barplot}
# create a unique code for each speaker's production of each phone

# create summary df to summarize the percent 
# of misalignment that fall in 10ms or 20ms
df_sum <- df_mfa %>%
  group_by(speaker) %>%
  summarise(n_phone=n(),
            n_phone_10=sum(diff < 0.01),
            n_phone_20=sum(diff < 0.02),
            per_10=(n_phone_10/n_phone)*100,
            per_20=(n_phone_20/n_phone)*100,
            res=res,
            speech_rate=mean(speech_rate)) %>%
  ungroup() %>%
  distinct() %>%
  pivot_longer(cols=starts_with("per_"),
               names_to = "range",
               names_prefix = "per_",
               values_to = "percent",
               values_drop_na = TRUE)

df_sum$speaker<-gsub("ma_","",as.character(df_sum$speaker))

# plot the percentages
p_percentage <- ggplot(df_sum, aes(x=res, y=percent, fill=res, shade=speaker)) + 
  #scale_fill_manual(values = c("dodgerblue1", "darkgoldenrod2", 
  #                              "coral1", "coral2",
  #                              "dodgerblue2","darkgoldenrod2",
  #                              "forestgreen")) +
  facet_grid(~range) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  geom_bar(stat='identity', width = 0.8, position = position_dodge(width = 0.9), alpha=0.8) +
  labs(x = "Variation of Mandarin", y = "Percent of aligner boundaries \n within threshold of human placement") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 100))

ggsave(p_percentage, filename = "asa_graphs/percentage.png",  width = 9, height = 5, bg = "transparent")

```

## Modelling

1) Testing only within the MFA vs. Suyuan comparisons: significant variation by variety? Significant variation by speaker?

```{r model}
df$diff_log <- log(df$diff)
df$speech_rate_log_s <- scale(log(df$speech_rate))

df_mfa <- filter(df, other_type=="mfa")

library(optimx)

# attempt to a linear mixed-effect model..?
mod_ML <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df_mfa,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_ML_0 <- lmer(diff_log ~ speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df_mfa,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_REML <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df_mfa,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=T
            )

mod_REML_0 <- lmer(diff_log ~ res + speech_rate_log_s +
                     (0 + speech_rate_log_s | speaker),
            data=df_mfa,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=T
            )

# mod_REMLPH <- lmer(diff_log ~ res + speech_rate_log_s + phone_category +
#               (1 + speech_rate_log_s + phone_category | speaker),
#             data=df_mfa,
#             control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
#             REML=T
#             )
# 
# mod_REMLPH_0 <- lmer(diff_log ~ res + speech_rate_log_s + 
#                      (1 + speech_rate_log_s + phone_category | speaker),
#             data=df_mfa,
#             control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
#             REML=T
#             )


summary(mod_ML)
anova(mod_ML, mod_ML_0, test="Chisq") # no significant overall effect of variety
anova(mod_REML, mod_REML_0, test="Chisq", refit=F) # significant variation across participants
#anova(mod_REMLPH, mod_REMLPH_0, test="Chisq", refit=F) # significant variation across phones?
```

2) Testing across the MFA vs. Suyuan and Sijia vs. Suyuan comparisons: significant difference in accuracy?

```{r model}
mod_comp_ML <- lmer(diff_log ~ other_type + speech_rate_log_s + res +
              (1 + other_type | speaker),
            data=df,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_comp_ML_0 <- lmer(diff_log ~ speech_rate_log_s + res + 
              (1 + other_type | speaker),
            data=df,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

summary(mod_comp_ML) #
# model predictions:
# 1) mean onset displacement for Beijing for manual alignments
exp(fixef(mod_comp_ML)[1] + fixef(mod_comp_ML)[2] + fixef(mod_comp_ML)[3] * mean(df$speech_rate_log_s)) # ~6 ms
# 2) same for manual vs. mfa
exp(fixef(mod_comp_ML)[1] + fixef(mod_comp_ML)[3] * mean(df$speech_rate_log_s)) # ~16 ms

anova(mod_comp_ML, mod_comp_ML_0, test="Chisq") # significant difference between manual alignment consistency vs. consistency between manual and mfa
```

## Closer look at sounds that behave differently across Mandarin varieties
Beijing, Tianjin: sh/zh/ch
Southern Manarin: drop nasals: m/n/ng
Beijinng Tianjin: drop stops and h: p/t/k/b/d/g/h
```{r}
phone_set <- c("sh", "zh", "ch", "m", "n", "ng", "p","t","k","b","d","g","h")
df1 <- subset(df_mfa,
  simple_phone %in% phone_set
)

ggplot(df1, aes(y=diff*1000, x=res, color=speaker)) +
  #scale_color_manual(values = c("dodgerblue1", "darkgoldenrod2", 
  #                              "coral1", "coral2",
  #                              "dodgerblue2","darkgoldenrod2",
  #                              "forestgreen")) +
  facet_wrap(~phone_category) +
  geom_boxplot() +
  scale_y_log10() +
  theme_bw() + 
  labs(x = "Variation of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none")
```

```{r}
df1$diff_log <- log(df1$diff)
df1$speech_rate_log_s <- scale(log(df1$speech_rate))

# attempt to a linear mixed-effect model..?
mod_ML_1 <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df1,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_ML_0_1 <- lmer(diff_log ~ speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df1,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

anova(mod_ML_1, mod_ML_0_1, test="Chisq") # no significant overall effect of variety
```

```{r}
study <- c("Raymond et al. \n (2002)", "McAuliffe et al.\n (2017)","McAuliffe et al.\n (2017)", "MacKenzie & Turton \n (2020)", "MacKenzie & Turton \n (2020)")
fa <- c("human", "MFA", "FAVE", "DARLA", "FAVE")
displacement <- c(17, 17, 19.3, 12.4, 14.3)
prior_study <- data.frame(study, fa, displacement)

prior_study$study_f = factor(prior_study$study, levels=c("Raymond et al. \n (2002)", "McAuliffe et al.\n (2017)","MacKenzie & Turton \n (2020)"))

p_study <- ggplot(prior_study, aes(x=fa, label=displacement, y=displacement, fill=fa)) +
  geom_col() +
  geom_text(position = position_dodge2(width = 0.9, preserve = "single"), vjust=2, color="white") +
  facet_grid(~study_f, 
             scales = "free_x",
             space = "free_x",
             switch = "x") +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme(strip.placement = "outside",                      
        strip.background = element_rect(fill = "white"),  
        strip.text=element_text(size=14, colour="black"),
        #axis.title = element_blank(),
        axis.text=element_text(size=12, colour="black"),
        axis.line = element_line(colour = "gray"),
        axis.title=element_text(size=16, face="bold",),
        panel.background = element_blank()) +
    theme(legend.position = "none") +
  ylab("Onset Displacement \n w/ Human Alignment (ms)")

ggsave(p_study, filename = "prior_study.png",  width = 9, height = 5, bg = "transparent")
```
