---
title: "MFA_analysis_2"
author: "Suyuan Liu"
date: '2022-10-24'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(data.table)
library(lme4)
library(ggh4x)

# prepare for levenshtein mapping
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
source("levenshtein.r")
```

# compare suyuan and sijia alignment
```{r load_data}
df_manual <- read.csv("/Users/liusuyuan/Desktop/cleaned_MFA/df_manual.csv", header=TRUE)
```

```{r data_overview}
df_manual %>%
  group_by(res, speaker) %>%
  summarise(sd=sd(diff, na.rm=T),
            mean=mean(diff, na.rm=T),
            median=median(diff, na.rm=T),
            under_20_ms=mean(diff < 0.02, na.rm=T),
            over_100_ms=mean(diff > 0.1, na.rm=T)) %>%
  ungroup()

df_manual %>%
  summarise(sd=sd(diff, na.rm=T),
            mean=mean(diff, na.rm=T),
            median=median(diff, na.rm=T),
            under_20_ms=mean(diff < 0.02, na.rm=T),
            over_100_ms=mean(diff > 0.1, na.rm=T))
```

```{r manual comparison by speaker}
##By speaker###
df_means_manual_speaker <- df_manual %>%
  group_by(speaker, res) %>%
  summarise(diff_mean=mean(diff),
            diff_median=median(diff),
            mean_symbol="μ",
            diff_mean_label=paste0(round(diff_mean*1000), " ms"),
            diff_median_label=paste0(round(diff_median*1000), " ms")) %>%
  ungroup()

mean_fun <- function(x){
  print(x)
  return(data.frame(y = x, label = paste0("  ", x, " ms")))
}

ggplot(df_manual, aes(y=diff*1000, x=res, color=res, shade=speaker)) +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=res), col=NA, alpha=0.15) +
  geom_text(data=df_means_manual_speaker, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9)) +
  geom_text(data=df_means_manual_speaker, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9)) +
  geom_point(data=df_means_manual_speaker, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9),
             shape=21, fill=NA) +
  geom_point(data=df_means_manual_speaker, aes(y=diff_median*1000), size=1, position=position_dodge(width=0.9),
             shape=16, fill=NA) +
  geom_text(data=df_means_manual_speaker, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9)) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000),
                limits=c(0.01, 1000)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank())
```

```{r manual comparison by variety}
df_means_manual <- df_manual %>%
  group_by(res) %>%
  summarise(diff_mean=mean(diff),
            diff_median=median(diff),
            mean_symbol="μ",
            diff_mean_label=paste0(round(diff_mean*1000), " ms"),
            diff_median_label=paste0(round(diff_median*1000), " ms")) %>%
  ungroup()

mean_fun <- function(x){
  print(x)
  return(data.frame(y = x, label = paste0("  ", x, " ms")))
}

ggplot(df_manual, aes(y=diff*1000, x=res, color=res)) +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=res), col=NA, alpha=0.15) +
  geom_text(data=df_means_manual, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9)) +
  geom_text(data=df_means_manual, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9)) +
  geom_point(data=df_means_manual, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9),
             shape=21, fill=NA) +
  geom_point(data=df_means_manual, aes(y=diff_median*1000), size=1, position=position_dodge(width=0.9),
             shape=16, fill=NA) +
  geom_text(data=df_means_manual, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9)) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10, 30, 100, 300, 1000, 3000),
                limits=c(0.01, 1000)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank())

```

```{r model}
df_manual$diff_log <- log(df_manual$diff)
df_manual$speech_rate_log_s <- scale(log(df_manual$speech_rate))

library(optimx)

mod_ML <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df_manual,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_ML_0 <- lmer(diff_log ~ speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df_manual,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_REML <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df_manual,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=T
            )

mod_REML_0 <- lmer(diff_log ~ res + speech_rate_log_s +
                     (0 + speech_rate_log_s | speaker),
            data=df_manual,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=T
            )

#mod_REMLPH <- lmer(diff_log ~ res + speech_rate_log_s + phone_category +
#              (1 + speech_rate_log_s | speaker),
#            data=df_manual,
#            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
#            REML=T
#            )

#mod_REMLPH_0 <- lmer(diff_log ~ res + speech_rate_log_s + phone_category +
#                     (0 + speech_rate_log_s | speaker),
#            data=df_manual,
#            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
#            REML=T
#            )


summary(mod_ML)
anova(mod_ML, mod_ML_0, test="Chisq") # no significant overall effect of variety
anova(mod_REML, mod_REML_0, test="Chisq", refit=F) # significant variation across participants
#anova(mod_REMLPH, mod_REMLPH_0, test="Chisq", refit=F)
```


# compare mfa2 & suyuan

```{r load_data}
df <- read.csv("/Users/liusuyuan/Desktop/cleaned_MFA/df_model.csv", header=TRUE)
```

```{r data_overview}
df %>%
  group_by(res, speaker) %>%
  summarise(sd=sd(diff, na.rm=T),
            mean=mean(diff, na.rm=T),
            median=median(diff, na.rm=T),
            under_20_ms=mean(diff < 0.02, na.rm=T),
            over_100_ms=mean(diff > 0.1, na.rm=T)) %>%
  ungroup()

df %>%
  summarise(sd=sd(diff, na.rm=T),
            mean=mean(diff, na.rm=T),
            median=median(diff, na.rm=T),
            under_20_ms=mean(diff < 0.02, na.rm=T),
            over_100_ms=mean(diff > 0.1, na.rm=T))
```

```{r by_speaker}
df_means_speaker <- df %>%
  group_by(speaker, res) %>%
  summarise(diff_mean=mean(diff),
            diff_median=median(diff),
            mean_symbol="μ",
            diff_mean_label=paste0(round(diff_mean*1000), " ms"),
            diff_median_label=paste0(round(diff_median*1000), " ms")) %>%
  ungroup()

mean_fun <- function(x){
  print(x)
  return(data.frame(y = x, label = paste0("  ", x, " ms")))
}

ggplot(df, aes(y=diff*1000, x=res, color=res, shade=speaker)) +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=res), col=NA, alpha=0.15) +
  geom_text(data=df_means_speaker, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9)) +
  geom_text(data=df_means_speaker, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9)) +
  geom_point(data=df_means_speaker, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9),
             shape=21, fill=NA) +
  geom_point(data=df_means_speaker, aes(y=diff_median*1000), size=1, position=position_dodge(width=0.9),
             shape=16, fill=NA) +
  geom_text(data=df_means_speaker, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9)) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000, 3000),
                limits=c(0.5, 3500)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n") +
  theme(legend.position = "none") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank())
```

```{r by_variety}
df_means <- df %>%
  group_by(res) %>%
  summarise(diff_mean=mean(diff),
            diff_median=median(diff),
            mean_symbol="μ",
            diff_mean_label=paste0(round(diff_mean*1000), " ms"),
            diff_median_label=paste0(round(diff_median*1000), " ms")) %>%
  ungroup()

mean_fun <- function(x){
  print(x)
  return(data.frame(y = x, label = paste0("  ", x, " ms")))
}

ggplot(df, aes(y=diff*1000, x=res, color=res)) +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=res), col=NA, alpha=0.15) +
  #geom_text(data=df_means, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9)) +
  #geom_text(data=df_means, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9)) +
  geom_point(data=df_means, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9),
             shape=21, fill=NA) +
  geom_point(data=df_means, aes(y=diff_median*1000), size=1, position=position_dodge(width=0.9),
             shape=16, fill=NA) +
  geom_text(data=df_means, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9)) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000, 3000),
                limits=c(0.5, 3500)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n",
       title = "Human-Aligner Agreement") +
  theme(legend.position = "none") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank(),
        plot.title = element_text(size = 20, face = "bold"))
  #geom_hline(yintercept=17.8, linetype="dashed", color = "red") +
  #geom_hline(yintercept=6.7, linetype="dashed", color = "red")

#ggsave("/Users/liusuyuan/Documents/1_School/1_UBC/2_2021Spring/MFA_project/manuscript/plots/human-aligner.pdf", width = 8, height = 5)
```

```{r model}
df$diff_log <- log(df$diff)
df$speech_rate_log_s <- scale(log(df$speech_rate))

library(optimx)

mod_ML <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_ML_0 <- lmer(diff_log ~ speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=F
            )

mod_REML <- lmer(diff_log ~ res + speech_rate_log_s +
              (1 + speech_rate_log_s | speaker),
            data=df,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=T
            )

mod_REML_0 <- lmer(diff_log ~ res + speech_rate_log_s +
                     (0 + speech_rate_log_s | speaker),
            data=df,
            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
            REML=T
            )

#mod_REMLPH <- lmer(diff_log ~ res + speech_rate_log_s + phone_category +
#              (1 + speech_rate_log_s | speaker),
#            data=df_mfa,
#            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
#            REML=T
#            )

#mod_REMLPH_0 <- lmer(diff_log ~ res + speech_rate_log_s + phone_category +
#                     (0 + speech_rate_log_s | speaker),
#            data=df_mfa,
#            control=lmerControl(optimizer = "optimx", calc.derivs = FALSE, optCtrl = list(method = "nlminb", starttests = FALSE, kkt = FALSE)),
#            REML=T
#            )


summary(mod_ML)
anova(mod_ML, mod_ML_0, test="Chisq") # no significant overall effect of variety
anova(mod_REML, mod_REML_0, test="Chisq", refit=F) # significant variation across participants
#anova(mod_REMLPH, mod_REMLPH_0, test="Chisq", refit=F)
```


# plot human-aligner and human-human in one plot
```{r by_variety}
df$method = "Human-Aligner"
df_manual$method = "Human-Human"
df_combined = bind_rows(df, df_manual)
df_means$method = "Human-Aligner"
df_means_manual$method = "Human-Human"
df_combined_means = bind_rows(df_means, df_means_manual)

ggplot(df_combined, aes(y=diff*1000, x=res, color=method, shape=method)) +
  # using violins to show distributions per vowel
  geom_violin(aes(fill=method), col=NA, alpha=0.15) +
  geom_point(data=df_combined_means, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9), fill=NA) +
  geom_point(data=df_combined_means, aes(y=diff_median*1000, fill = method), size=1, position=position_dodge(width=0.9)) +
  scale_shape_manual(values=c(21, 24)) +
  geom_text(data=df_combined_means, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9),
            show_guide = FALSE) +
  geom_text(data=df_combined_means, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9),
            show_guide = FALSE) +
  geom_text(data=df_combined_means, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9),
            show_guide = FALSE) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000, 3000),
                limits=c(0.1, 3500)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n",
       color = "Comparison",
       fill = "Comparison",
       shape = "Comparison") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank(),
        strip.text = element_text(size=14, face="bold"),
        legend.position = "bottom",
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))

#ggsave("/Users/liusuyuan/Documents/1_School/1_UBC/2_2021Spring/MFA_project/manuscript/plots/by_variety.png", width = 7, height = 5)

```

```{r by_speaker}
df$method = "Human-Aligner"
df_manual$method = "Human-Human"
df_combined = bind_rows(df, df_manual)
df_means_speaker$method = "Human-Aligner"
df_means_manual_speaker$method = "Human-Human"
df_combined_means_speaker = bind_rows(df_means_speaker, df_means_manual_speaker)

ggplot(df_combined, aes(y=diff*1000, x=res, color=method, shape=method, shade=speaker)) +
  # using violins to show distributions per vowel
  facet_nested_wrap(~res + speaker, nrow = 2, ncol = 6, scales="free_x") +
  geom_violin(aes(fill=method), col=NA, alpha=0.15) +
  geom_text(data=df_combined_means_speaker, aes(y=diff_mean*1000, label=mean_symbol), size=4, position=position_dodge(width=0.9),
            show_guide = FALSE) +
  geom_text(data=df_combined_means_speaker, aes(y=diff_mean*1000*1.7, label=diff_mean_label), size=4, position=position_dodge(width=0.9),
            show_guide = FALSE) +
  geom_point(data=df_combined_means_speaker, aes(y=diff_median*1000), size=2.5, position=position_dodge(width=0.9)) +
  geom_point(data=df_combined_means_speaker, aes(y=diff_median*1000, fill=method), size=1, position=position_dodge(width=0.9)) +
    scale_shape_manual(values=c(21, 24)) +
  geom_text(data=df_combined_means_speaker, aes(y=diff_median*1000*0.59, label=diff_median_label), size=4, position=position_dodge(width=0.9),
            show_guide = FALSE) +
  # adding the median separately as a point
  scale_y_log10(breaks=c(1, 3, 10,30, 100, 300, 1000, 3000),
                limits=c(0.1, 3500)) +
  scale_fill_brewer(palette="Dark2") +
  scale_colour_brewer(palette="Dark2") +
  theme_bw() + 
  labs(x = "Variety of Mandarin\n", y = "Onset displacement (ms) \n",
       color = "Comparison",
       fill = "Comparison",
       shape = "Comparison") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        panel.grid=element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size=12),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 14, face = "bold"))

#ggsave("/Users/liusuyuan/Documents/1_School/1_UBC/2_2021Spring/MFA_project/manuscript/plots/by_speaker.png", width = 10, height = 6)

```


# plot participant metadata
```{r participant_location_map}
library(sf)
library(mapchina)

df <- china %>%
        filter(Code_Province %in% as.character(10:82))
sf_use_s2(FALSE)
df <- df %>%
        group_by(Name_Province) %>%
        summarise(geometry = st_union(geometry))

df$participant_n = case_when(df$Name_Province == "上海市" ~ 1,
                             df$Name_Province == "云南省" ~ 1,
                             df$Name_Province == "北京市" ~ 1,
                             df$Name_Province == "天津市" ~ 1,
                             df$Name_Province == "江苏省" ~ 1,
                             df$Name_Province == "广东省" ~ 1,
                             df$Name_Province == "四川省" ~ 1,
                             TRUE ~ 0)
df$Name_Province_eng = case_when(df$Name_Province == "上海市" ~ "Shanghai",
                             df$Name_Province == "云南省" ~ "Xinan",
                             df$Name_Province == "北京市" ~ "Beijing",
                             df$Name_Province == "天津市" ~ "Tianjin",
                             df$Name_Province == "江苏省" ~ "Nanjing",
                             df$Name_Province == "广东省" ~ "Guangdong",
                             TRUE ~ NA)

df_training$training_n = case_when(
  grepl("河口", df_training$Name_County) ~ 1,
  df_training$Name_Province == "北京市" ~ 1,
  grepl("武汉", df_training$Name_County) ~ 1,
  TRUE ~ 0)

df_training$Name_Province_eng_training = case_when(
  #grepl("河口", df_training$Nam) ~ "Hekou",
  df_training$Name_Province == "北京市" ~ "Beijing",
  grepl("武汉", df_training$Name_County) ~ "Wuhan",
  TRUE ~ NA)

ggplot() +
  geom_sf(data = df, aes(fill = participant_n)) +
  scale_fill_distiller(palette = "Greys", 
                       direction = 1,
                       limits = c(0, 0.5),
                       guide = guide_legend(),
                       #breaks = c(0,1,3,5,7),
                       name = "Participant number") +        
  theme_bw() +
  #ggtitle("Mandarin varieties distribution") +
  theme(plot.title = element_text(color="black", size=14, face="bold.italic"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none") +
  ggrepel::geom_label_repel(
    data= df,
    aes(label = Name_Province_eng, 
        geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0,
    #segment.color = 'transparent',
    segment.color = NA,
    size = 3,
    fill = alpha(c("white"),0.5)) +
  ggrepel::geom_label_repel(
    data= df_training,
    aes(label = Name_Province_eng_training, 
        geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0,
    #segment.color = 'transparent',
    segment.color = "grey",
    size = 3,
    fill = alpha(c("red"),0.5))  

#ggsave("/Users/liusuyuan/Documents/1_School/1_UBC/2_2021Spring/MFA_project/manuscript/plots/Mandarin_varieties.png", width = 6, height = 4)
    
```


# Correlation between human-human and human-aligner
r=0; there is no relation between the variable.
r=+1; perfectly positively correlated.
r=-1; perfectly negatively correlated.
r= 0 to 0.30; negligible correlation.
r=0.30 to 0.50; moderate correlation.
r=0.50 to 1 highly correlated.

```{r}
colnames(df_manual)[which(names(df_manual) == "phone_start_suyuan")] <- "phone_start_manual_2"
colnames(df_manual)[which(names(df_manual) == "phone_start_sijia")] <- "phone_start_manual"
colnames(df_manual)[which(names(df_manual) == "diff")] <- "Human-Human"
colnames(df)[which(names(df) == "diff")] <- "Human-Aligner"


df_combined_wider <- merge(df_manual, df, by = "phone_start_manual")
```

```{r}
df_combined_means_speaker_wider <- 
df_combined_means_speaker[, c("method", "diff_mean", "speaker", "res")] %>%
  pivot_wider(names_from = method, values_from = diff_mean)

#df_combined_means_speaker <- 
df_combined[, c("method", "diff", "speaker", "res", "word_id", "phone", "word",
                "previous", "following")] %>%
  pivot_wider(names_from = method, values_from = diff)

#Pearson correlation score: 0.1567868
cor(df_combined_means_speaker_wider$`Human-Aligner`,
    df_combined_means_speaker_wider$`Human-Human`, 
    method = 'pearson')

df_combined_means_variety_wider <- 
df_combined_means[, c("method", "diff_mean", "res")] %>%
  pivot_wider(names_from = method, values_from = diff_mean)

#Pearson correlation score: 0.2934998
cor(df_combined_means_variety_wider$`Human-Aligner`,
    df_combined_means_variety_wider$`Human-Human`, 
    method = 'pearson')


#Pearson correlation score: 0.05423127
cor.test(df_combined_wider$`Human-Aligner`,
    df_combined_wider$`Human-Human`, 
    method = 'pearson')

library(ggpubr)

scaleFUN <- function(x) sprintf("%.2f", x)


ggplot(df_combined_wider, aes(x=`Human-Aligner`*100, 
                              y=`Human-Human`*100))+
  geom_point(aes(color=res.x), alpha=0.1) +
  geom_smooth(color = "black", method="lm") +
  theme_bw() + 
  scale_color_brewer(palette = "Dark2") +
  ylab("Human-Human Onset Displacement\n(Log-scaled, in ms)") +
  xlab("Human-Aligner Onset Displacement\n(Log-scaled, in ms)") +
  labs(color = "Mandarin Variety") +
  #stat_cor(method="spearman", size = 5) +
  scale_x_continuous(trans = "log", labels = scaleFUN) +
  scale_y_continuous(trans = "log", labels = scaleFUN) +
  annotate("label", x = 0.01, y = 0.005, label = "r = 0.054, p < 0.05") +
  theme(axis.title=element_text(size=16, face="bold",),
        axis.text=element_text(size=14, colour="black"),
        legend.position = "none")

#ggsave("/Users/liusuyuan/Documents/1_School/1_UBC/2_2021Spring/MFA_project/manuscript/plots/displacement_cor.png", width = 8, height = 5)
```

